Mendelian randomization for causal inference in genome-wide association studies

### 2022-03-01 mr-review.Rmd
added text about MR-GxE

>>>>

@iong_latent_2020 build on other research using mixture models in the two-sample MR setting with GWAS summary 
statistics. Like MR-Clust, MR-PATH, as the method from @iong_latent_2020 is called, works with the Wald ratio 
estimates from multiple SNP instruments. Coining the term "mechanistic heterogeneity" to refer to the 
distribution of Wald ratio estimates due to SNP involvement in distinct causal pathways, @iong_latent_2020 
specify a $J$-component normal mixture model. They use an expectation maximization algorithm to fit the model 
parameters (the component weights and the parameters for each of the $J$ normal distributions) and suggest a 
Bayesian information criterion-based approach to choosing $J$ [@schwarz1978estimating].

@wang_weakinstrument_2021 propose causal effect test statistics that are robust under weak instrument asymptotics 
[@staiger1994instrumental]. Specifically, they extend three econometric methods to two-sample MR with GWAS 
summary statistics [@anderson1949estimation;@kleibergen_pivotal_2002;@moreira_conditional_2003] and study the 
theoretical properties and the performance in applications. @wang_weakinstrument_2021 extend their methods to 
demonstrate conditions under which their estimators are equivalent to those in works from @zhao_statistical_2020, @windmeijer_use_2019, and @bowden_improving_2019.

@spiller_detecting_2018 and @spiller_interaction-based_2020 present a method, MR-GxE, that exploits gene by 
environment interactions to detect and to adjust for bias due to horizontal pleiotropy. They leverage the insight 
that a SNP instrument-covariate interaction that induces differences in SNP instrument-exposure associations 
@spiller_interaction-based_2020 treat a SNP-covariate interaction as an instrumental variable, and, by so doing, they impose assumptions on that interaction. Together, these assumptions require that MR-GxE assume that horizontal pleiotropy effects are constant across the study cohort. The authors devise a three-step inferential procedure for two-sample MR with GWAS summary statistics:

1. Estimate the SNP-exposure and SNP-outcome associations at a range of covariate values  
1. Regress the SNP-outcome estimated associations on the SNP-exposure estimated associations  
1. Treat the resulting slope as the causal effect estimate and the intercept as the mean horizontal pleiotropy effect.

@spiller_interaction-based_2020 point out that MR-GxE can be used as a sensitivity analysis tool to choose a set 
of valid SNP instruments from a collection of SNPs in settings where there may be a violation of the constant 
pleiotropy assumption.
